Polynomial Time and Dependent Types

### 2023-10-25 paper/paper.tex
[ paper ] more little fixes

There are at least two ways that a functional programming language can
be seen as representing polynomial time, differing in how the size of
the input is measured. One approach is to define the input as a closed
expression, including what we might think of as its input, and
computation of the result is polynomial time in the combined size. A
second approach is that the input is ``externally'' provided, so we
consider open terms with a free variable representing the input. So a
judgement $x : \tyNat \vdash M : A$ declares a program that computes
results of type $A$ in time polynomial in the size of the natural
number $x$. We take this latter approach in this paper.

@@

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
  Thanks to Anton Lorenzen and Fredrik Nordvall Forsberg for their
  comments on an earlier revision, and to the anonymous POPL reviewers
  for their detailed and interesting reviews. This work was funded by
  the \grantsponsor{EPSRC}{Engineering and Physical Sciences Research
    Council}{https://www.ukri.org/about-us/epsrc/}. Grant number
  \grantnum{EPSRC}{EP/T026960/1}, \emph{AISEC: AI Secure and
    Explainable by Construction}.
  % FIXME: add Fred, when his comments have been included.
\end{acks}

>>>>

There are at least two ways that a functional programming language can
be seen as representing polynomial time, differing in how the size of
the problem to be computed is measured. One approach is to consider
closed expressions, combining the program with its input, and
computation of the result is polynomial time in the combined size. A
second approach is that the input is ``externally'' provided, where we
consider open terms with a free variable representing the input. So a
judgement $x : \tyNat \vdash M : A$ declares a program that computes
results of type $A$ in time polynomial in the size of the natural
number $x$. We take this latter approach in this paper.

@@

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
  Thanks to Anton Lorenzen and Fredrik Nordvall Forsberg for their
  comments on an earlier revision, and to the anonymous POPL reviewers
  for their detailed and interesting reviews. This work was funded by
  the \grantsponsor{EPSRC}{Engineering and Physical Sciences Research
    Council}{https://www.ukri.org/about-us/epsrc/}. Grant number
  \grantnum{EPSRC}{EP/T026960/1}, \emph{AISEC: AI Secure and
    Explainable by Construction}.
\end{acks}
